### 2017-05-24 Paper.tex
2017.5.24

%% Paper.tex
%% V1.4b
%% 2017/05/17
%% by jyc
%% See:
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{User-level High-performance flexible \\ network traffic processing library}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Yanchao Jing}
\IEEEauthorblockA{School of Shanghai\\
Jiao Tong University\\
Email:13541333146@163.com}
\and
\IEEEauthorblockN{Guangtao Xue}
\IEEEauthorblockA{School of Shanghai\\
Jiao Tong University\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{ShiyouQian}
\IEEEauthorblockA{School of Shanghai\\
Jiao Tong University\\
Email:}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The abstract goes here.
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
Nowadays the ever-increasing demand for quality of service (Qos) and network
security including policy-based routing,firewall and virtual private network(VPN)
\cite{IEEE:1},In order to satisfy the requirement,An in-depth
understanding of the Internet traffic profile is needed.Accordingly, the deep packet inspection(DPI)
become a crucial hot technology\cite{IEEE:2}.In a nutshell,a DPI system Firstly has to collect packets
from the network interface cards(NIC),in order to create a data structure to represent the incoming packets
as network flows and then to forward or store the received packets for further processing.So it is extremely
important to capture all exchanged packets between end-systems on the monitored link in very high speed links
(e.g. multiple Gbps).\cite{IEEE:2}.many DPI application increasely high-level entities and constructs such as
TCP flows Udp infomation HTTP headers and so on.Therefore a high-performance universal network traffic capture
library is required to develop DPI high-level applications expediently.
\newline\indent Recently,For high-performance,enormous relative thoeritical methods and tools have been proposed such as
zero-copy technology, multi-core capture,multi-thread method and FPGA tools and so on. FPGA tools is a optimization 
method of add-on cards\cite{IEEE:3} which is not in our focus scope.This paper mainly focus on software optimization;
the zero-copy and multi-core technology can help avoid sparse sampling\cite{IEEE:4}.A famous tool to finish these optimization
is PF\_RING\cite{IEEE:5}.PF\_RING kernel module runs as a software interrupt handler that stores incoming packets to a memeory-mapped
buffer,shared with user-level stub. there are also many multi-thread proposed methods that can improve the entire performance.
the core function of multi-thread is to  make the best of multi-core performance.Although there are many kinds of theoretical
technologies, there is hardly a flexible high-performance network traffic library. In this situation, developing a strong DPI application
becomes a huge challenge.
\newline\indent To overcome this diffculty, in this paper we present a network traffic library(noff),a high-performance passive network
traffic monitoring framework.noff provides the high-level functionality needed by monitoring application and provide a quite flexible
framework to adjust some functions.Application developers can easily create magic high-level applications based on noff.
\newline\indent To acquire aggressive performance optimization, we use a flexible sur-level multi-thread structure towards a  
Multi queue NIC or just plain NIC.It can make best of the multi-core performance to achieve a high throughput.
\newline\indent To reduce the overhead of unneccessary memory copying.The noff regard the PF\_RING as underlying kernel component. the subzero packet copy
of pf\_ring can contribute to reduce the total overhead and noff use google tcmalloc[] framework to reduce the malloc frequency.
\newline\indent Apparently, In order to synchronic the multi-thread system, mutex is a directly solution.But lock can bring implicative
performance loss. noff make use of intel concurrency queue to make the entire system lockfree and get higher performance.
\newline\indent To accommodate heavy loads,noff introduces the notion of first input first output(FIFO) stream management algorithm.Under heavy load,traditional
monitoring systems usually drop arriving packets in a random way,severely affecting any following stream reassembly process[].noff always
drop the stream which is firstly built connection in hash table.this method can relatively reduce the impact from dropping packets.
\newline\indent Noff also provides a flexible and expressive Application programming interface that allows programmers to conifgure many kinds of
stream aspect capture process procedure.Our design introduces two novel features:(\uppercase\expandafter{\romannumeral1})noff use C++11 bind and function
features to design callback interfaces between two modules.Programmers can flexiblely add new modules or change modules just in main config
function rather than in module body.and(\uppercase\expandafter{\romannumeral2})it offers more control for tolerating packet loss
under high load throughput. And noff also support single queue NIC multi-thread scaling and multi queue based on RSS(receive-side scaling)
transparent parallelization of stream processing simultaneously.
\newline\indent We have evaluated noff in a 10GbE environment using real compus traffic.and showed that it outperforms existing alternatives like Libnids
and Snort's stream5 in variety of scopes.For example,our results demonstrate that noff can capture and deliver all streams with lower CPU utilization.
The main contributions of this paper are:
\newline 1.We propose the noff,which is a universal network traffic library.It provides a flexible and aggressive module framework to offer programmers a easy-use platform
to develop network plug-in componentents.
\newline 2.We introduce some applied software techniques to optimize traditional products and approaches.
through these techniques,noff can deliver transport-layer streams for two times higher traffic rates
than previous approaches.
\newline 3.We introduce first in first out(FIFO),a algorithm that enables graceful adaptation to overload
conditions by dropping packets of old streams,and favoring packets that belong to recent and shorter streams.

\section{software Feature}
The design of noff fouces attention on two key objectives: structure flexbility and runtime performance.In this part,we introduce the main aspects of noff across these two dimensions.

% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
\subsection{Flexble Module Framework and Interface}

noff adopts a novel callback module interface which is based on C++11 bind and function feature.noff is composed of serveral modules.
these modules can be divided into serveral parts by function or logical structure. the logical structure includes
capturer,dispatcher,ip,tcp,udp,app and so on,which is correspond to Tcp/Ip network model to some extent.each layer will offer one
or more callback interface to many upper layer modules.programmers can add their module based on lower layer module without change any
existing code.At the same time,noff can use different modules at a same layer to adapt to different demand.

\subsection{Pf\_ring Patch}
Pf\_ring is a new type of network socket that dramatically improves the packet capture speed.It use User-space
User-space ZC (new generation DNA, Direct NIC Access) drivers for extreme packet capture/transmission speed as the 
NIC NPU (Network Process Unit) is pushing/getting packets to/from userland without any kernel intervention. 
noff use Pf\_ring as the capture driver and arrive aggressive runtime speed.

\subsection{First Input First Output}
noff introduces First Input First Output(FIFO) stream algorithm enable the system to invest its resources effectively
during overload. This algorithm can resolve the problem that sudden traffic bursts or overload conditions may force the packet
capturing subsystem to fill up its buffers and randomly drop packets in a haphazard manner.noff maintains a connection timer queue which 
sort the stream based on the first connection time in ascending order.when the stream buffer is full. noff would drop the stream in the head of
timer queue.As long as the percentage of used memory is below a user-defines threshold,FIFO drops no packets.In this way,the newly stream will be 
accommodated with higher probability.Programmers can also force to maintain some priority streams by update the timestamp.

\subsection{Flexible Stream reassembly}
Libnids accomplished its Tcp Reassemble machine by simplify the linux kernel code. However,the callback interface of libnids is poor and
it has no ability to cope with SYN Flood attack and connection abnormity because of none timer. Noff add a passive keep-alive timer to close the dead abnormal connection when
a stream connection has no data interaction during a long time. Noff also abandon the tcp buffer to reduce memeory copy. Noff reserves nearly all of the function in libnids.

\subsection{Memory Allocations Optimization}
Noff uses google Tcmalloc to optimize the memeory copy performance. The Tcmalloc implementation takes approximately 50 nanoseconds for the same operation pair when ptmalloc2 takes approximately 300.
The Tcmalloc assigns each thread a thread-local cache. Small allocations are satisfied from the thread-local cache. 
Objects are moved from central data structures into a thread-local cache as needed, and periodic garbage collections are used to migrate memory back from a thread-local cache into the central data structures.

\subsection{Parallel Processing}
Noff offers two kinds of multi-thread platforms which inherent support for multi-core systems.the two kinds are all
achieved by transparently creating a number of worker threads for

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

  \bibitem{IEEE:1}
  Liu, Duo, et al. "High-performance packet classification algorithm for many-core and multithreaded network processor." International Conference on Compilers, Architecture and Synthesis for Embedded Systems ACM, 2006:334-344.

  \bibitem{IEEE:2}
  Antonello, Rafael, et al. "Deep packet inspection tools and techniques in commodity platforms: Challenges and trends." Journal of Network \& Computer Applications 35.6(2012):1863-1878.

  \bibitem{IEEE:3}
  Qiao, Siyi, et al. "Network recorder and player: FPGA-based network traffic capture and replay." International Conference on Field-Programmable Technology IEEE, 2015.

  \bibitem{IEEE:4}
  Ali, Sardar, et al. "On mitigating sampling-induced accuracy loss in traffic anomaly detection systems." ACM SIGCOMM Computer Communication Review 40.3(2010):4-16.

  \bibitem{IEEE:5}
  PF\_RING. http://www.ntop.org/products/packet-capture/pf\_ring/
\end{thebibliography}

% that's all folks
\end{document}